Push-Pull Signal-Function Functional Reactive Programming

### 2012-11-30 Docs/IFL2012/introduction.tex
more intro

An FRP program is generally evaluated by supplying a behavior or signal function
to an evaluation function, which supplies input (in the case of a signal
function) and actuates its outputs as they are produced. For our weather system
example, the evaluation function is responsible for sampling the sensors,
receiving notification from the lightning strike detectors and pushing the
occurrences to the signal function, and drawing the picture which is the output
signal of the signal function.

Classic FRP was first described as a system for interactive animations~\cite{Elliott1997}.
Recent work on classic FRP has focused on efficient implementation. One approach
to efficiency is to separate the evaluation of behaviors and events, since
as was discussed above, the suitable strategy for evaluating signals is not the
same as that for events. The initial implementations of FRP made use of
pull-based evaluation for both behaviors and events.
``Reactive''~\cite{Elliott2009}, as well as more recent systems such as
``reactive-banana''~\cite{Apfelmus}, make use of push-based evaluation for
events and pull-based evaluation for behaviors. This is known as ``push-pull''
evaluation.

@@

A recent extension of signal-function FRP called N-Ary FRP~\cite{Sculthorpe2011}
describes a method of typing signal functions which, as we will show, enables
the push-based evaluation of events in a signal-function FRP system. The notion
of signal vectors allows the representation of signal function inputs and
outputs as combinations of signals and events, rather than a single signal which
may contain multiple values, including option values for events. Signal vectors
are uninhabited types, which can be used to describe partial or full
representations of the signal function inputs and outputs.

TimeFlies,\footnote{The sentence ``Time flies like an arrow.'' is a 
favorite quotation of one of the author's philosophy instructors, used to
demonstrate the ambiguity of language. The origin of the quotation is unknown.}
is a push-pull signal-function FRP system, implemented in the Haskell
programming language. Our contributions with this system are as follows:

\begin{itemize}
  \item We show that typing signal-function FRP with signal vectors allows
        push-pull evaluation of signal functions.
  \item We demonstrate a programmable evaluation interface that allows the user
        of TimeFlies to take advantage of this behavior while interfacing
        TimeFlies to a wide range of IO systems.
  \item We give benchmark results which demonstrate that the push-pull
        evaluation strategy makes TimeFlies more efficient than the current
        pull-based implementation of signal-function FRP.
\end{itemize}

>>>>

An FRP program is generally evaluated by supplying a behavior or signal function
to an evaluation function, which provides input (in the case of a signal
function) and actuates its outputs as they are produced. For our weather system
example, the evaluation function is responsible for sampling the sensors,
receiving notification from the lightning strike detectors and pushing the
occurrences to the signal function, and drawing the picture which is the output
signal of the signal function.

Classic FRP was first described as a system for interactive animations~\cite{Elliott1997}.
%% Recent work on classic FRP has focused on efficient implementation. One approach
%% to efficiency is to separate the evaluation of behaviors and events, since
%% as was discussed above, the suitable strategy for evaluating signals is not the
%% same as that for events. 
\rn{Let's not be non-specific: ``focus on efficient implementation.
  One approach...'', if there aren't actually other examples of
  ``efficient implementation'' and we're really just talking about push-pull.}
%%
The initial implementations of FRP made use of
pull-based evaluation for both behaviors and events.
``Reactive''~\cite{Elliott2009}, as well as more recent systems such as
``reactive-banana''~\cite{Apfelmus}, make use of push-based evaluation for
events and pull-based evaluation for behaviors. This is known as ``push-pull''
evaluation \new{and is asymptotically more efficient}.

@@

A recent extension of signal-function FRP called N-Ary FRP~\cite{Sculthorpe2011}
describes a method of typing signal functions which, as we will show, enables
the push-based evaluation of events in a signal-function FRP system. The notion
of {\em signal vectors} allows the \new{(type-level)} representation of signal-function inputs and
outputs as combinations of signals and events, rather than a single
signal \new{of composite} values, including option values for events. Signal vectors
are uninhabited types, which can be used to describe partial or full
representations of the signal function inputs and outputs.
\rn{But what's the limitation of N-Ary FRP?! It doesn't allow full
  push-pull right?  Or wait, the SV technique ``enables'' push-pull,
  but the previous sculthorpe implementation just didn't do it?  Clarify.}

\new{In this paper we introduce}
TimeFlies,\footnote{The sentence ``Time flies like an arrow.'' is a 
favorite quotation of one of the author's philosophy instructors, used to
demonstrate the ambiguity of language. The origin of the quotation is unknown.},
a push-pull signal-function FRP system, implemented in the Haskell
programming language. Our contributions are as follows:
% \rn{Should be the paper contributions, not the system!}

\begin{itemize}
  \item We show that typing signal-function FRP with signal vectors allows
        push-pull evaluation of signal functions. \rn{FIRST?}
  \item We demonstrate a programmable evaluation interface that allows the user
        of TimeFlies to take advantage of this behavior while interfacing
        TimeFlies to a wide range of IO systems.
  \item We give benchmark results which demonstrate that the push-pull
        evaluation strategy makes TimeFlies more efficient than the current
        pull-based implementation of signal-function FRP.
\end{itemize}
